---
title: "Chunks_and_Tabs"
author: "Adrienne Dunk"
date: "2022-10-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




# A New Section With Tabs {.tabset}

## Tab 1

Here's some text

## Tab 2

Here's some graphics:

```{r tab_graph 1, echo = FALSE}
require(palmerpenguins)
boxplot(
  flipper_length_mm ~ species, data = penguins,
  main = "Flipper Length and Species",
  ylab = "Flipper Length (mm)"
)

```

## Tab 3

Here's a numbered list:

1. Item 1
1. Item 2
1. Item 3
1. **Item 4** (in bold)

# New 1st Level header (not there are no tabs)


## Subsection 1

This section is not in a tab

## Subsection 2

Still no tabs!

# Tab Section 2 {.tabset .tabset-pills}

This seciton uses a 'pill style' of tabs!

## Tab 1

Here's some text

## Tab 2

Here's some graphics:

```{r tab_graph 2, echo = FALSE}
boxplot(
  flipper_length_mm ~ species, data = penguins,
  main = "Flipper Length and Species",
  ylab = "Flipper Length (mm)"
)

```

## Tab 3

Here's a numbered list:

1. Item 1
1. Item 2
1. Item 3
1. **Item 4** (in bold)