---
title: "Frequentist Concepts"
author: "Adrienne Dunk"
subtitle: Analysis of Environmental Data
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: 
      collapsed: false
      smooth_scroll: false
---


# Q1: Binomial Probability 1
What is the probability of observing a count of exactly 3 success in a binomial distribution with n = 4, and p=0.75?

```{r}
dbinom(3, 4, 0.75) #42%

```

---


# Q2: Binomial Probability 2
What is the probability of observing a count of 3 success or fewer in a binomial distribution with n=4, and p=0.75?

```{r}
pbinom(3, 4, 0.75) #68%

```

---

# Q3: Binomial Probability 3
What is the probability of observing more than 3 successes in a binomial distribution with parameters n=5 and p=0.75.
```{r}
pbinom(3, 5, 0.75) #probability of getting 3 or fewer (37%)

1-pbinom(3, 5, 0.75) #probability of getting >3 (63%)

```

---

# Q4: Normal Probability 1

What is the probability of observing a value of less than 1.2 from a normally distributed population with mean = 2 and standard deviation = 2?

```{r}
pnorm(1.2, mean = 2, sd = 2) #34%


```

---

# Q5: Normal Probability 2

What is the probability of observing a value greater than 1.2 from a normally-distributed population with mean = 2 and standard deviation = 2?

```{r}
1-pnorm(1.2, mean = 2, sd = 2) #66%

```

---

# Q6: Normal Probability 3
What is the probability of observing a value between 1.2 and 3.2 from a normally distributed population with mean = 2 and standard deviation = 2?
```{r}
a <- pnorm(3.2, mean = 2, sd = 2) #prob less than 3.2
b <- pnorm(1.2, mean = 2, sd = 2) #prob less than 1.2
between <- a-b
between #38%

#pnorm(3.2, mean = 2, sd = 2)- pnorm(1.2, mean = 2, sd = 2) ##Not Run - same answer as above, less syntax

```

---

# Q7: Central Limit Theorem 1
With additional sampling, the shape of the histogram better approximates the shape of the skewed distribution.

---

# Q8: Central Limit Theorem 2
With additional sampling, the shape of the histogram approached a normal distribution.


# Q9-11: Central Limit Theorem 3
### 9
The histogram was roughly normal after the first sample, but continued to normalize around the sample mean with repeated sampling

### 10
The drastic change in the shape of the sampling distribution when it was increased from 1 to 15 is due to the calculation of the sample mean. When you only take 1 sample, the sample mean is the sampled value; however, when you take 15 samples, the sample mean is calculated from that sample distribution. This gives a better approximation of the mean around which the sample values are distributed.

### 11
The two primary factors that determine the width of the sampling distribution of the mean are: sample variance and sample size.

# Q12: Library of Babel 1

```{r}
words3letters <- 25^3
words3letters #15,625

```


# Q13: Library of Babel 2
```{r}
B <- 410 * 40 * 80

B2 <- 410 * 40 *81

B2/B

#1.0125B Books (1.10125*(2*10^1,834,097))

```








