---
title: "Using Models 1"
author: "Adrienne Dunk"
subtitle: Analysis of Environmental Data
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: 
      collapsed: false
      smooth_scroll: false
---


# load data

```{r Import Data}
require(here) #load here into session
here() #check directory

catrate <- read.csv(here("data", "catrate.csv"))
head(catrate)
summary(catrate)

```

# Q1 Catastrophic Rate Histogram
Create a histogram of salamander reproduction catastrophic rates. 
```{r}
hist(catrate$cat.rate, main = "Reproduction Catastrophic Rates", xlab = "Catastrophic Rate")

```

---


# Q2-4: Normality Test

### Q2: Conduct a Shapiro-Wilk test for the salamander catastrophic rates.
```{r}
shapiro.test(catrate$cat.rate)
#p-value of 0.04

```
### Q3: What is the null for the Shapiro Test
The null value for the Shapiro Test is that the population is normally distributed.

### Q4: Are the catastrophic rates normally-distributed?
No, the low p-value indicates that we should reject the null. These rates are not sampled from a normally distributed population.

---

# Q5-7: T-Test 1
### Q5: Run the t-test that the catastrphic rate is different from the pond late-refilling rate.
```{r}
t.test(catrate$cat.rate, mu = 2/7)

```

### Q6: State the null in plain English
The null hypothesis is that the average of the catastrophe rate and the average of late-refilling ponds are statistically the same.

### Q7: Is this a one- or two-tailed test
This is a two-tailed test because we only asked if they were the same or different and we did not specify a direction (greater/less than).

---

# Q8-10: T-Test 2

### Q8: What is the p-value from the Q5 t-test?
The p-value is 0.01 which means that the catastrophe rate average and the pond late-refilling rates are not the same. With a p-value of 0.01, we have 1% chance of saying that they are different when they are actually the same.

### Q9: What are the confidence intervals from the Q5 t-test?
The 95% confidence intervals were 0.353 to 0.726. The interval does not include 0.

### Q10: Did you reject the null? If so, why?
Yes, I reject the null that the mean of the catastrophe rates are equal to the rates of the late-refilling rates. I reject it because the null is less than the pre-determined rejection rate of 0.05. Also the confidence interval does not contain 0 indicating that there is sa true difference.

---

# Q11-13 Wilcoxon Test 1

### Q11: Conduct a one-sample Wilcoxon rank sum test that the catastrphic rate is different than the pond late-refilling rate.

```{r}
wilcox.test(catrate$cat.rate, mu = 2/7)
```
### Q12: Compare the Wilcox and t-test p-values
The Wilcox p-value is 0.006 while the t-test p-value was 0.019. In both cases, we would reject the null that these two rates are the same; however, the Wilcox test provides a lower chance of having a false-positive as the p-value is an order of magnitude smaller.

### Q13: Did you reject the null, if so, why?
I rejected the null because the p-value of the Wilcox test was below my rejection rate of 0.05.

---

# Q14-15: Test Comparison

### Q14: Compare the conclusions you could draw from the t-test and Wilcox test.
Both tests allow you to draw the same conclusion that the pond late-refilling rate is different than the catastrphic rate. 

### Q15: Which test was more appropriate for these data?
The Wilcox test was more appropriate for these data because the catastrophic rates were not normally distributed.

---

# Q16-17: Flippers - Normality Tests

### Q16: conduct tests of normality for the flipper lengths of Chinstrap and Adelie penguins.
```{r}
require(palmerpenguins)

penguin_dat <- droplevels(subset(penguins, species != "Gentoo"))
summary(penguin_dat)
dat_adelie <- subset(penguin_dat, species == "Adelie")
dat_chinstrap <- subset(penguin_dat, species == "Chinstrap")
shapiro.test(dat_adelie$flipper_length_mm) #is normal
shapiro.test(dat_chinstrap$flipper_length_mm) #is normal

```
### Q17: Interpret the test results.
The Shapiro-Wilk test indicates that the Adelie and Chinstrap flipper lengths are each normally distributed. The p-values are 0.72 and 0.81 respectively. Given this, we cannot reject the null hypthesis that the sample populations are normally distributed.

---

# Q18: Flippers - Histograms
Create a single figure of two side-by-side histograms of the flipper lengths of the penguin species.

```{r}
require(here)
png(filename = here("Penguin_Flippers.png"), width = 800, height = 600)

par(mfrow = c(1,2))
hist(dat_adelie$flipper_length_mm,
     main = "Adelie",
     xlab = "Flipper Length (mm)")
hist(dat_chinstrap$flipper_length_mm,
     main = "Chinstrap", 
     xlab = "Flipper Length (mm)")

dev.off


```
---

# Q19-20: Flippers T-test

Conduct a two-sample t-test of the alternative hypothesis that the Adelie have different flipper lengths than the Chinstrap penguins.
### Q19: 
The alternative hypothesis of the test is that the average flipper length of Adelie penguins is different than the average flipper length of Chinstrap penguins. This is a two-tailed test because we are not hypothesizing about the direction of any possible difference.


### Q20:
```{r}
t.test(penguin_dat$flipper_length_mm ~ penguin_dat$species, alternative = c("two.sided"))


```


